Data Recovery and Pulse Position Modulation with a Photon Number Resolving SNSPD

### 2023-10-05 manuscript.tex
overleaf update

\begin{abstract*} 
Photon number resolution is an emerging capability of advanced Superconducting Nanowire Single Photon Detectors. If leveraged to it's full potential, PNR capability can have a profound impact on the usefulness of SNSPDs is certain quantum applications including linear optical quantum computing, quantum networks, and quantum sensing. Discrimination of not just the number of photons in an optical pulse but also pulse arrival time with high accuracy is an open problem, complicated by the nuanced way in which these two degrees of freedom are intertwined in the response function of these detectors. In this work we put a differential readout SNSPD to the ultimate test. We test it's capabilities in a Pulse Position Modulation experiment whereby data is sent in the arrival time of optical pulses derived from a 20~GHz optical clock. We show the detector is capable of discriminating the arrival time of photons to 50~ps wide bins with high accuracy while simultaneously providing photon number information about the impinging optical pulses. We find that a careful statistical analysis of the PNR response is necessary to back-out a precise measurement of pulse arrival time.
\end{abstract*}

@@

\textcolor{red}{ The following is very rough, taken from years ago when I first started writing the manuscript }

\protect\hypertarget{..ux2fcodeux2ftest_1}{}{}

This study aimed to evaluate the feasibility of transmitting high clock-rate pulse position modulated (PPM) data using a mode-locked laser and receiving it with a low jitter superconducting nanowire single-photon detector (SNSPD). The investigation was driven by recent advancements in NbN SNSPDs, which have achieved a jitter as low as 50 ps at the FW(1/100)M level, enabling the demonstration of PPM with 50 ps slot widths and a 20 GHz clock. The aim was to increase the data rate by a factor of 10, from 2 GHz to 20 GHz, in the next generation of the Deep Space Optical Communication (DSOC) project.

This study aimed to evaluate the feasibility of transmitting high clock-rate pulse position modulated (PPM) data using a mode-locked laser and receiving it with a low jitter superconducting nanowire single-photon detector (SNSPD). The investigation was driven by recent advancements in NbN SNSPDs, which have achieved a jitter as low as 50 ps at the FW(1/100)M level, enabling the demonstration of PPM with 50 ps slot widths and a 20 GHz clock. The aim was to increase the data rate by a factor of 10, from 2 GHz to 20 GHz, in the next generation of the Deep Space Optical Communication (DSOC) project.

\hypertarget{pulse-position-modulation-for-deep-space-communication}{%
\subsection{Pulse Position Modulation for Deep Space Communication}\label{pulse-position-modulation-for-deep-space-communication}}

Deep Space Optical Communication has been a growing field of study in recent years, as researchers look for ways to communicate with spacecraft that are far away from Earth. In this article, we will focus on the use of Pulse Position Modulation (PPM) for deep space optical communication.

While other modulation techniques such as quadrature amplitude modulation (QAM) have been used in the past, it has been shown that in the photon-starved regime, PPM is the best approach for sending data. This is because there is not enough light to measure the phase of the signal.

The main challenge in deep space optical communication is the high loss and distance that the link must traverse. This limits the communication from the spacecraft, as it is limited by the power available on the spacecraft. This means that the communication protocol is limited by the number of bits that can be sent per unit of energy on the spacecraft, also known as photon information efficiency or bits per photon.

@@

>>>>

\begin{abstract*} 
Superconducting Nanowire Single Photon Detectors are a type of time-correlated photon detector with low jitter performance especially in the mid-infrared. They are useful for classical communication over high loss channels --such as across deep space-- and for quantum communication for which signals are restricted to the few-photon level. For classical communication, high photon information efficiency communication may be achieved with Pulse Position Modulation (PPM) whereby data is encoded in the arrival time of an optical pulse with respect to a clock. In the process of demonstrating PPM on a 20~Ghz clock, we study the effects of Photon Number Resolution (PNR) in new low-jitter types of SNSPDs. These PNR effects complicate fixed-threshold triggering of RF pulses from the SNSPD, and corrupt arrival time measurements if not properly managed. We demonstrate methods for simultaneous arrival time and photon number measurement which enables high clock rate PPM for space applications as well as high rate quantum communication and computing applications that benefit from photon number resolution.
\end{abstract*}

@@

Deep Space Optical Communication has been a growing field of study in recent years, as researchers look for ways to communicate with spacecraft that are far away from Earth. The main challenge in deep space optical communication is the high loss and distance that the link must traverse. This limits the communication from the spacecraft, as it is limited by the power available on the spacecraft. This means that the communication protocol is limited by the number of bits that can be sent per unit of energy on the spacecraft, also known as photon information efficiency or bits per photon.

In this article, we will focus on the use of Pulse Position Modulation (PPM) for deep space optical communication. PPM can achieve higher photon information efficiency than coherent detection techniques {[}cite a method here, like is QAM one of them? See dolinar paper{]} for which the need to measure phase of the incoming optical signal limits minimum transmission power.

PPM involves sending an optical pulse in one of \(2^M\) possible time slots. The transmissio of each such pulse carries M bits of information.

For larger M, more data may be sent with a single optical pulse, thereby allowing a spacecrat to that is severely power limited to send more data over a high loss optical channel back to earth. The Deep Space Optical Communicaiton project managed by the Jet Propulsion Laboratory (JPL) aims to demonstrate optical communication using PPM with the Psyche spacecraft from distances of \todo[inline]{XX to XX} Au.

@@

This study aims to evaluate the feasibility of transmitting high clock-rate pulse position modulated (PPM) data using a mode-locked laser and receiving it with a low jitter superconducting nanowire single-photon detector (SNSPD). The investigation was driven by recent advancements in NbN SNSPDs, which have achieved a jitter as low as 50 ps at the FW(1/100)M level, enabling the demonstration of PPM with 50 ps slot widths and a 20 GHz clock. The aim was to increase the data rate by a factor of 10, from 2 GHz to 20 GHz, in the next generation of the Deep Space Optical Communication (DSOC) project.

\hypertarget{section}{%
\section{--------------------}\label{section}}
